CK, here is my next round of comments (to your comments of May 30/93).

p. 2, 1: Fine.

p. 3, 1: I removed the claim that our method is more efficient.

p. 4, 1: I have added a paragraph at the end of Section 3.

p. 5, 1: I have added a footnote to the proof, to address the comment
	of Referee 1.

p. 6, 1 a) unnecessary
	b) made the change to Euclidean plane with a line at infinity

p. 7, 1: done
      2: unnecessary: you are worried that we include it 
	so that we can be sure when degenerate cases will occur
	but the condition of this lemma is exactly the one defined
	in Defn 4.2 as the condition of nondegeneracy.
	Therefore, I have removed Lemma 4.3 and added a defn of $S_i$	
	to the second line of Section 5. 

p. 8, 1: done

p. 9, 1: I have removed (a) the rest of the paragraph on p. 9, 
	starting at `Once the first intersection conic'
			(b) the footnote on p. 10
		    and (c) step 7 of the algorithm on p. 11.

	I have changed the statement of step 4 in the algorithm, 
	but there are still some missing details.
	Could you please fill these in?
	Also, we should reconsider whether we want to make the change
	at all, since the step 4 is becoming very bulky.

p. 10, 	1: footnote removed.
	2: changed the wording of the paragraph about focal length
	
p. 11,  1: I don't understand?
	2: Removed step 7.
	3: Done.
	4: I added a remark about minor axis length to end of 4th paragraph
	   of Section 4.3.
	   Then I referred to the above computations in step 6 of
	   the algorithm.

p. 14, 1: I have replaced the entire proof (since it is needed in Thm 7.1).
	I have not cited your geometric proof in the thesis because
	if they find out that a geometric proof already exists, there
	is little argument not to simply include that proof here
	(at the request of Referee 3).
	I will tell Rossignac that a geometric proof would be much longer.

p. 16, 1: OK.

p. 18: added a footnote to `we will provide a new and very short algebraic
	proof'; changed order of sentences; changed 
	`can simplify previous algebraic proofs' to 
	`can simplify previous proofs' (is this OK?)

p. 20, 1: Instead, I referred directly to your thesis in the footnote
	to step 4 of the algorithm.

p. 22, 1: I did not change it back, because: 
	we use $\alpha_1 + \alpha_2$ to find the appropriate table
	in Table 3 to search, *then* we use $cos \theta$ to search within
	the table, not $cos (\alpha_1 + \alpha_2)$.




	





