\documentstyle[11pt,letterhead]{bletter}
\signature{Prof. John K. Johnstone\\jj@cs.jhu.edu}
\begin{document}
\begin{letter}
{Attn: Shirley Foatz(?)\\
Springer Verlag\\
Journal Production (Visual Computer)\\
P.O. Box 10 52 80\\
D-6900 Heidelberg\\
Fed. Rep. Germany
}

\opening{Dear Ms. Foatz(?):}

I am writing concerning corrections to the proofs of the paper
`A localized method for intersecting plane algebraic curve segments'
by John Johnstone and Michael Goodrich, which is manuscript number 324
for Visual Computer.
I spoke to you about this today on the phone.
(Please pardon the misspelling of your name: I am transcribing from a
note of the secretary.)
I am sending the marked proofs, the original manuscript, and the purple
copy of the figures to you
by normal mail, along with the copyright transfer statement.
However, so that you can correct the proofs immediately, I am also
outlining below all of my corrections.
(I am keeping the extra copy of the proofs, the photos, and the original
figures, since I understand you don't need those anymore.)
We get 50 offprints free of charge, and this is all that we want.
I am faxing the Copyright transfer statement as well in case it is urgent.

The following changes to the proofs are necessary:
\begin{enumerate}
\item
	page 1: there should be a double star (rather than a single star)
	after the name of the second author
\item
	page 2, column 1, Example 2.1, line 2: remove the hyphen
	so that `two-plane' becomes `two plane'.
\item
	page 2, column 2, line 3: `is discarded' should be `are discarded'
\item
	page 4, column 1, line 5: `(of coordinated crawl)' should be
	`(of the coordinated crawl)'
\item
	page 4, column 2, three lines above Lemma 3.1: `Theorem 3.1'
	should be `Lemma 3.1'
\item
	page 4, column 2, line 5 of Lemma 3.1: `which have' should be
	`which has'
\item
	page 5, column 1, three lines above section 3.2: `due the expense'
	should be `due to the expense'
\item
	page 5, column 2, equation: rather than splitting the equation
	after y(A), the second line should start with `x(Pactive) ...'.
	Thus the first line of the equation would be
	`$((y(A') < y(C')) \neq (y(A) < y(C)))$'
	and the second line would be \\
	`$\vee x(P_{\mbox{\footnotesize{active}}}) 
	\geq \mbox{min} \{x(B),x(D)\}$'
\item
	page 6, column 2, equation: similarly, this equation should be
	split after the epsilon symbol so that the first line would be 
	`$|y(P_{\mbox{active}}) - y(P_{\mbox{dormant}})| < \epsilon$'
	and the second line would be
	`$\vee \ x(P_{\mbox{\footnotesize{active}}})>\mbox{min} \{x(B),x(D)\}$'
\item
	page 8, col 2, line 13: `curvce-curve' should be `curve-curve'
\item
	page 8, col 2, last line: `the' should be `then'
\item
	page 9, figure 10: the numbers 2 and 3 should be interchanged
	so that the order of the lines from left to right is 1,3,4,2.
\item
	page 10, col 1, line 4 of proof: `the length of AB' should be 
	`the length of the curve segment AB'
\item
	page 12, col 1, line 2: `exploding degree sets with' should
	be changed to `degree explosion is encountered in'
\item
	page 12, Garrity-Warren reference: the page numbers are 137-153
	(and it is in Volume 6, Number 2)
\item
	page 12, Johnstone-Bajaj reference: the paper `Sorting points along
	an algebraic curve' is in Volume 19, Number 5, pages 925-967
	of the October 1990 volume of SIAM Journal on Computing.
\end{enumerate}


If you have any questions, please do not hesitate to call me at
(301) 338-5560 or fax me at (301) 338-6134.

\closing{Sincerely, }
\end{letter}
\end{document}
