\documentstyle[11pt,letterhead]{bletter}
\signature{John K. Johnstone}
\begin{document}
\begin{letter}
{Crystal Norris\\
Senior Production Editor\\
SIAM\\
3600 University City Science Center\\
Philadelphia, PA  19104-2688}

\opening{Dear Ms. Norris:}

Here is the draft of the paper `Sorting Points along an Algebraic Curve' 
with editions.
I have made all of the corrections outlined on the edited copy (also enclosed),
except for those enumerated below.

% One of the main things that bothers me about this typesetting of the
% paper is the constantly varying spacing between lines, caused by the
% presence of superscripted hats on many symbols (particularly arc symbols).
% This certainly detracts from the appearance of the paper.
% With double-spacing, the spacing between lines does not vary, but
% I suppose that it is probably illegal to double-space the whole thing.

\begin{enumerate}
\item
Because our local system does not have the appropriate fonts,
some of the paper is missing: the titles of the figures (e.g., `Fig. 1'),
and the names of the authors in the bibliography.
Cory Gray says that this is not a problem, because you can run my 
Latex file yourself with the appropriate fonts.
\item
I think that my wording on p. 12 is more grammatical and more understandable:
`a segment of the curve that one might wrongly worry has been cut out of the curve'.
\item
I believe `Singularities' in title of Section 5.1 {\em should} be capitalized.
\item
I prefer the use of a comma on p. 22 of original: `as well as points that touch
	 ---, face ---'.
\item
`Semi-cubical parabola' (p. 43) is the spelling used in the literature.
\item
I am unsure about some of the information in the following references: 1,3,16.
\item
There is no publisher given in Allerton conference proceedings, so I can't specify one.
\item
Capitalization is not allowed in your bibliography package.
For example, it should be GANITH in reference 9, but it insists on printing Ganith.
It is very important that reference 26 have a capital O for `O(nloglogn)' since
a small `o' has a totally different meaning.
This also applies to Reference 25.
I don't know how to correct this.
\end{enumerate}

I apologize for the delay in getting this to you. 
The file required a great deal of Latex manipulation (`hacking'),
compounded by the fact that I had to be at a conference in Detroit this week.

Please feel free to call me if you have any questions, at (301) 338-5560.

\closing{Sincerely,}
\end{letter}
\end{document}






