Dear Prof. Goldman:

	It was good to see you at the conference.
As we discussed, C.-K. and I have a method for the plane/natural-quadric
intersection problem that you might be interested in.
It comes out of our work on the natural quadric intersection problem 
and, like your method, uses Dandelin spheres, but in a different way.
For simplicity and conciseness, I am sending you only the heart of the method: 
computing the plane/cone intersection, ellipse case.
The hyperbola case is identical
to the ellipse and can be collapsed into the ellipse case.
The simple special cases, such as sphere and cylinder, parabola, etc., 
are each very short and left out:
if you are interested in seeing those, I would be glad to send them.
As you can see, once the weaponry for computing the focal length is developed
(Section 2.1), the rest of the computation is quite short.

If you feel that these ideas are interesting, and have not submitted your 
version of this problem, we would be happy to collaborate on this paper.
Please let me know what you think.

There are three accompanying figures: 
would you please send me your fax number so that I can send them to you?

I am sending the same material to Prof. Miller, and also sending you both
a copy of Dandelin's paper by U.S. mail.

Finally, I have a favour to ask of you.   Our department is looking for senior
faculty.  If you know of any senior faculty in the solid modeling/geometry
community that might be interested, and if you could spread the word amongst
your colleagues, I would really appreciate it.

			All of the best, 	
			John
