\documentstyle[11pt,letterhead]{bletter}
\signature{John K. Johnstone\\jj@cs.jhu.edu}
\begin{document}
\begin{letter}
{Prof. Jose L. Encarnacao\\
Editor-in-Chief, Computers and Graphics\\
Fachgebiet Graphisch-Interaktive Systeme\\
Technische Hochschule Darmstadt\\
Wilhelminenstr. 7\\
D-6100 Darmstadt\\
FRG
}

\opening{Dear Prof. Encarnacao:}

I enclose the copyright agreement and keyword page for the article
`Computing the Intersection of a Plane and a Natural Quadric' by
John K. Johnstone and C.-K. Shene.
I also enclose copies of all figures.
There was a request for a change to Figure 3,
and I would like to change Figure 2.
The other figures have not changed, but I enclose copies so that you have
a complete list.
Finally, we have become aware of some related work that deserves mention.
Therefore, on page 1, at the end of paragraph 4, we would like to add the
following sentence:

\begin{quote}
An algorithm for intersecting a plane and natural quadric that is quite
similar has recently been developed by Miller and Goldman \cite{MiGo91}.
\end{quote}

\noindent and add the following citation:

\begin{quote}
James R. Miller and Ronald N. Goldman, Using Tangent Balls
to Find Plane Sections of Natural Quadric Surfaces, Draft, April 1991.
\end{quote}

\noindent Since this added citation causes a renumbering of other citations,
I have enclosed a copy of the paper with this addition.
Thank you very much for your attention to this matter.

Please do not hesitate to call me at (410) 516-5560 if you have any questions.

\closing{Sincerely,}
\end{letter}
\end{document}

