\documentstyle{article}
\pagestyle{empty}
\sloppy      
          
%        The layout of the page
%
\marginparwidth 0pt
\oddsidemargin  0pt
\evensidemargin  0pt
\marginparsep 0pt
\leftmargin -0.25in

\topmargin   -0.6in  %0pt

\textwidth   7.0in
\textheight  9.0in

\renewcommand{\baselinestretch}{1.0}


\begin{document}


\parskip=4pt
\parindent=0pt

\begin{center}
{\Large\bf MSI Workshop on}

\vspace{0.5cm}
 
{\LARGE\bf COMPUTATIONAL GEOMETRY }

\end{center}

\vspace{0.2cm}

\begin{center}
{\Large\bf
October 25--26, 1991 \\[0.2cm]

SUNY Stony Brook \\
Stony Brook, NY \\
}

\vspace{1.5cm}

{\bf Sponsored by the U.S. Army Research Office, and the\\ Mathematical
Sciences Institute, Stony Brook.}

\end{center}

\begin{center} 
{\bf Organizers:}
\end{center}

{\bf
\begin{tabular}{ll}

Joseph S. B. Mitchell & \hskip1.8in Steven S. Skiena\\
Applied Math & \hskip1.8in Computer Science \\
SUNY Stony Brook & \hskip1.8in SUNY Stony Brook\\
Stony Brook, NY 11794 & \hskip1.8in Stony Brook, NY 11794\\
(516) 632-8366 & \hskip1.8in (516) 632-9026/8470\\
jsbm@ams.sunysb.edu & \hskip1.8in skiena@sbcs.sunysb.edu\\

\end{tabular}
}

\begin{center}

{\large\bf Conference Information}

\end{center}

{\bf Location:\ \ } The workshop will be held in room S240 of the
Mathematics Tower at the State University of New York at Stony Brook.
There will be signs posted directing you to the workshop.  The
enclosed Campus Map and Travel Guide gives information about how one
gets to SUNY Stony Brook.  (For flyers, we recommend the
Islip/MacArthur airport.)  Airport limos from JFK or LaGuardia cost
about \$35 (each way) and will provide door-to-door service, {\em
but should be reserved in advance}.  (The limo driver will need
your flight info, etc, in order to arrange to meet you at the airport.)
Limo companies include Spartan Limo ((516) 928-5454; prices are
\$28 to, \$32 from JFK or LaGuardia, \$17 to/from Islip; they accept
cash or check), and Winston (toll-free (800) 424-7767; prices are \$35
from JFK or LaGuardia, \$20 from Islip; they accept cash, credit
cards, and NY state checks).

\medskip
{\bf Parking:\ \ }
If you intend to have a car on campus (personal or rental), we need to
get a parking sticker for you.  Please send us the make, model, and
license plate of your personal car (or indicate that it will be a
rental), and we will either send you (time permitting) a parking pass
or will provide one immediately upon your arrival.  (Sending this
information by email or FAX would be ideal.)  A parking sticker will
permit you to park in almost any lot near the Math Tower.  If you
arrive without one, park as close as possible, and run in to the
Workshop Registration desk to receive a permit before you get
ticketed!  If worse comes to worse, you can park in the garage (what the
signs call ``Visitors Parking''), but they will charge you about \$6.

\medskip
{\bf Accommodations:\ \ }
Participants who need overnight accommodations should make reservations
on their own, as early as possible.  Please let us know if we can
assist in any way.  

The least expensive hotel in the immediate area is the Holiday Inn
Express, located at 3131 Nessconset Highway (Route~347), 1/2 mile east
of the intersection with Nicolls Road (Route~97).  The Holiday Inn is
quite nice and newly opened this year.  It includes free HBO and a
nice continental breakfast.  The rooms are large, modern, and
comfortable.  When placing reservations, mention the Computational
Geometry Workshop at Stony Brook, in order to get the university rate.
Prices are \$65.00 for a single or double.  (Additional person is \$8)
{\em We have a block of rooms held until October 19, after which rooms
are available only on a first-come basis.} Phone: (516) 471-8000 (Use
the local number rather than the 800 number, in order to get the
discount rate).
Also, local faculty and graduate students have offered space on their
floors for visiting graduate students.
If you are interested in this service, inform the organizers as soon
as possible.

A fancier place with lots of charm and atmosphere is the Three Village
Inn, at 150 Main Street in Stony Brook.  (Be careful --- there are
many ``Main Streets'' in the area!)  They have a full-service restaurant
and are located within easy walking distance to many shops and cafes.
Prices are \$80.00 for a single, \$85.00 for a double, if you mention
that you are with the Stony Brook conference.  Phone: (516) 751-0555.

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{center}
 
{\normalsize\bf Preliminary Program}

\end{center}
\parskip=4pt
\parindent=0pt

{\normalsize\bf Friday, October 25, 1991}

Coffee and donuts 9:30--10:00 am, Room S240, Math Tower, SUNY Stony Brook.

\vskip 10pt

\begin{description}
\item[10:00]
{\it Separation and Approximation of Polyhedra.}
Subhash Suri,
{\sf Bellcore}

\item[11:00]
{\it Approximate Point Matching.}
Paul Heffernan,
{\sf William Patterson College}

\item[11:20]
{\it Constructing the Convex Hull of a Partially Sorted Set of Points.}
Michael T. Goodrich,
{\sf Johns Hopkins University}

\item[11:40]
{\it A New Intersection Algorithm using Circle Decomposition.}
John Johnstone,
{\sf Johns Hopkins University}

\item[12:00--12:20]
{\it Proximity Searching in 2- and 3-Dimensions: An Implicit Approach
for the $L_\infty$ Metric.}
Hossam ElGindy,
{\sf McGill University}
\end{description}

{\bf Lunch Break 12:20--1:30 pm}  (Deli platter lunch will be provided)

\begin{description}

\item[1:30]
{\it TBA.}
David Mount,
{\sf University of Maryland}

\item[2:30]
{\it Open Problem Session; Coffee Break.}

\item[3:30]
{\it TBA. (Topic: Interface of Graphics and CG)}
David Dobkin,
{\sf Princeton University}

\item[4:30] 
{\it Adaptive Polygonization of Trimmed NURBS Surfaces.}
%Adaptive Faceting of
%Trimmed NURBS Surfaces Using Constrained Delaunay Triangulation.}
H. Nebi Gursoy,
{\sf  Intergraph Corporation}

\item[4:50]
{\it Efficient Parallel Algorithms for Minimum Link Path and Related Problems.}
V. Chandru,
{\sf Purdue University}

\item[5:10]
{\it Fast Triangulation via Empty Spheres.}
Isabel Beichl (with Francis Sullivan),
{\sf Computing and Applied Mathematics Laboratory, NIST}

\item[5:30--5:50]
{\it TBA.}
Wm. Randolph Franklin,  % (will send abstract)
{\sf Rensselaer Polytechnic}

\end{description}

{\bf  Dinner 6:00--8:00 pm} (small groups at local restaurants)

{\bf Reception 8:00 pm} (location TBA) 

\bigskip

{\normalsize\bf Saturday, October 26, 1991}

Coffee and donuts 9:30--10:00 am, Room S240, Math Tower, SUNY Stony Brook.

\vskip 10pt

\begin{description}
\item[10:00]
{\it Shape Matching Algorithms for Visual Recognition.}
Dan Huttenlocher,
{\sf Cornell University}

\item[11:00]
{\it On Reconstructing Polyhedra from Parallel Slices.}
Joseph O'Rourke,
{\sf Smith College}

\item[11:20]
{\it Guaranteed-Quality Meshes for Curved Surfaces.}
Paul Chew,
{\sf Cornell University}

\item[11:40]
{\it TBA (Topic: New Results in Triangulation/Mesh Generation).}
Scott Mitchell,
{\sf Cornell University}

\item[12:00]
{\it Coping with Inconsistencies: A New Approach to Produce Quality
Triangulations for the Finite Element Method.}
Elefterios Melissaratos,
{\sf Rutgers}
\end{description}

{\bf Lunch Break 12:20--1:30 pm} (Deli platter lunch will be provided)

\begin{description}

\item[1:30]
{\it TBA.}
Mike Karasick,
{\sf IBM}

\item[2:30]
{\it A Line Sweep Thinning Algorithm.}
T. Pavlidis,
{\sf SUNY Stony Brook}

\item[2:50]
{\it The Robot Localization Problem.}
Rajeev Motwani,
{\sf Stanford University}

\item[3:10--3:30]
{\it Shape from Diameter: On the Recovery of
a Polygon's Shape from its Diameter Function.}
Anil Rao, {\sf USC}

\item[3:30]
{\it Placing Guards in Polygons with Holes.}
Iliana Bjorling-Sachs,
{\sf Rutgers University}

\item[3:50]
{\it A Pseudo-Output-Sensitive Algorithm for CSG to B-rep Conversion.}
Maged Tawfik, 
{\sf 3D/Eye Inc.}

\item[4:10--5:50]
{\it Open. Possibly follow-up on open problem discussions.}

\end{description}


\end{document}
