Rida,
	I enclose a few diverse polishings and questions.
It seems that at this stage we need to evaluate what still needs to be done.
I think that we are very close, and it would be great to finish up soon.
The theory seems to be almost complete.
Like yours, my schedule is freeing a bit, with the tumult of the
beginning of classes and a move to a new system now over.
I am certainly willing to come up to IBM for a day if we need to do 
anything in person.


The additions are as follows:

(1) In Definition 3.3, I changed the definition of inflection point.

(2) A remark on computation after Theorem 3.1.

(3) An addition before Definition 2.3 of $C_{\bf q}$,
	to motivate future use of $C_{\bf q}$.

(4) A bit more discussion on related work to end of Introduction, p. 1.
	You may want to add other articles you know about.



And some questions.

(1) Is it unsafe to use the term `critical point'?  
I like its connotation, but it has a different defn in your Hierarchical
Segmentation paper.  Is the definition in that paper popular?
Perhaps it is alright to overlap definitions, since the critical point
is immediately identified with the self-intersection.

(2) I'm probably misunderstanding, but
would it be possible to prove that dist(q,r(u)) is not 
always differentiable (second half of Proposition 1.2)
without the use of the directional derivative?
That is, just using (24) and (25).
The directional derivative certainly remains necessary in the discussion
after Definition 1.3.

(3) Formula (49), is $r'' = \sigma' t - \sigma^2 \kappa n$
	in the derivation of $b''$?

(4) Formula (50-51): I cannot figure out how to derive this (from (34)).
	Same for formula (57).

(5) You had mentioned problems with self-intersections: were those based
	upon implementation of the resultant?



By the way, your observations in Example 3.2 are quite remarkable,
Who would have thought that circles of curvature are related in 
such a way to the self-intersection polynomial?!
Also very impressed with your figures.
I will have to get your drawing package.


						John